---
title: "Get records by year and taxon (data_paper borreguiles)"
author: "Antonio J. Perez-Luque (@ajpelu)"
date: "Sep 2014"
output:
  html_document:
    theme: united
    highlight: pygments
  md_document:
    variant: markdown_github
---
#### Goals 
This scripts creates a table with the number of records by taxon and year in the Borreguiles dataset 

```{r someMetadata, eval=FALSE}
# -----------------------------------------------------------
#############################################################################
#### SCRIPT: gbif_borreguiles_004_observ_taxon_year.R                    ####
#### VERSION: 2                                                          ####
#### FECHA: 04/11/2014                                                   ####
#### AUTOR: ajperez@ugr.es                                               ####
#############################################################################
# -----------------------------------------------------------
```

```{r packages, warning=FALSE, message=FALSE} 
# -----------------------------------------------------------
# load library
library('lubridate')
library('reshape2')

# -----------------------------------------------------------
```


```{r directory}
# -----------------------------------------------------------
## Directorio 
di <- '/Users/ajpelu/Dropbox/MS/MS_DP_Borreguiles/git/BorreguilesDP'

# read file
mydata <- read.table(file=paste(di, '/2_diccionarios/gbif2014_c2.txt', sep=''), header=TRUE, sep='\t')
# Ojo revisar el campo fecha este en formato dd/mm/yyyy 
# -----------------------------------------------------------
```

Now we need to extract the year from date field.  

```{r prepareData}
#--------------------------------------------------------------------
# Extract year 
mydata$date <- as.Date(mydata$FECHA, "%d/%m/%Y")
mydata$year <- year(mydata$date)
#--------------------------------------------------------------------
```

Then we aggregate the data of each taxon to get the counts of each taxon in each year. We get a table with three fields: taxon, year and count. 
```{r aggregate}
# -----------------------------------------------------------
# Aggregate by year and taxon. Number of records by taxon and year 
aux <- aggregate(mydata$TAXON, by=list(mydata$TAXON, mydata$year),FUN='length')
names(aux) <- c('TAXON', 'year', 'count')
```

Now we transform the output table. We want a column (field) by each sampling year. We use the function `dcast` from `lubridate` [package](http://cran.r-project.org/web/packages/lubridate/index.html). We also replace the NAs values with ''. 

```{r agrregatebyyear}
#--------------------------------------------------------------------
aux2 <- dcast(aux, TAXON ~ year , value.var = 'count' )
aux2[is.na(aux2)] <- ''
aux2  
#--------------------------------------------------------------------
```


```{r exportResults}
# -----------------------------------------------------------
### EXPORTAR DATOS
write.csv(aux2, file=paste(di, '/2_diccionarios/taxones_year.csv', sep=''), row.names=FALSE)
# -----------------------------------------------------------
```
